---
title: "Practice lab 4 assignment"
author: "Andria Grace"
date: "2023-10-10"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
#empty environment
rm(list=ls())

#Loading libraries
library(dplyr)
library(tidyverse)
library(readxl)
library(table1)
library(doBy)
library(writexl)


#folder definitions
read.directory<-"C:/Users/Andria Grace/Desktop/Semester1/Computaional stats and probabaility Module/"
#read.directory<-"C:/Users/3108313/OneDrive - Fresenius Medical Care/Desktop/JGR/Academics/Yeshiva/Fall 2023/Graduates/Module 3/"
write.directory<-"C:/Users/Andria Grace/Desktop/Semester1/Computaional stats and probabaility Module/cdc.xlsx"
file.name<-"Module 4 - df.xlsx"
```

#1)	What’s the probability of
#a.	rolling a 6 on a die
```{r}
p<-1/6
p
```
#b.	rolling three consecutive times a 6
```{r}
p<-1/6*1/6*1/6
p
```
#c.	rolling a 6, a 1 and a 6 again
```{r}
p<-1/6*1/6*1/6
p
```
#2)	Define categories of random variables and give at least 2 examples for each group. 
#a.	Continuous
```{r}
## When a random variable takes continuous variable(Takes infinite values) is called continuous random variables.
##example: Time, Temperature
```

#b.	Categorical
```{r}
##It contains a finite number of categories or a group. eg: Gender, Blood Type
```

#c.	Discrete
```{r}
##  When a random variable takes specific or distinct variables within the interval is called discrete random variable.
##example:Number of Cars, rolling a die
```

#3)	Normal distribution
#a.	Load the CDC dataset
#Read data
```{r}
source("http://www.openintro.org/stat/data/cdc.R")
View(cdc)
#Functions
SD_pop<-function(x,n){sqrt(sum( (x - mean(x) )^2)/(n))}
```
#b.	Calculate all index points for the empirical rule
```{r}
#calculate BMI
cdc$weight_kg<-cdc$weight/2.2
cdc$height_cm<-cdc$height * 2.5
cdc$height_m <-cdc$height_cm / 100
cdc$BMI <- cdc$weight_kg / cdc$height_m^2
df<-cdc
mean(df$BMI)
SD_pop((df$BMI),length(df$BMI))
c(round(mean(df$BMI)-1*SD_pop(df$BMI,length(df$BMI)),2), "to", mean(df$BMI)+1*SD_pop(df$BMI,length(df$BMI))) #68%
c(mean(df$BMI)-2*SD_pop(df$BMI,length(df$BMI)), "to", mean(df$BMI)+2*SD_pop(df$BMI,length(df$BMI))) #95%
c(mean(df$BMI)-3*SD_pop(df$BMI,length(df$BMI)), "to", mean(df$BMI)+3*SD_pop(df$BMI,length(df$BMI))) #99.7%
```
#c.	Plot the normal distribution for BMI as a histogram in R
```{r}
hist(df$BMI)
```
#d.	Indicate +/-1, 2 and 3 standard deviations in the histogram
```{r}
hist(df$BMI)
abline(v=mean(df$BMI)-1*SD_pop(df$BMI,length(df$BMI)),col="red",lty=2)
abline(v=mean(df$BMI)-2*SD_pop(df$BMI,length(df$BMI)),col="green",lty=4)
abline(v=mean(df$BMI)-3*SD_pop(df$BMI,length(df$BMI)),col="blue",lty=4)
abline(v=mean(df$BMI)+1*SD_pop(df$BMI,length(df$BMI)),col="red",lty=2)
abline(v=mean(df$BMI)+2*SD_pop(df$BMI,length(df$BMI)),col="green",lty=4)
abline(v=mean(df$BMI)+3*SD_pop(df$BMI,length(df$BMI)),col="blue",lty=4)
```
#e.	Calculate the above stated values, the z-scores and the probabilities
```{r}
z_scores=(df$BMI-mean(df$BMI))/SD_pop((df$BMI),length(df$BMI))
quantile(z_scores,probs = c(68,95,99.7)/100)
```


#probabilities
```{r}
pnorm(mean(df$BMI)+1*SD_pop(df$BMI,length(df$BMI)),mean=mean(df$BMI),sd=SD_pop(df$BMI,length(df$BMI)), lower.tail = TRUE)-pnorm(mean(df$BMI)-1*SD_pop(df$BMI,length(df$BMI)),mean=mean(df$BMI),sd=SD_pop(df$BMI,length(df$BMI)), lower.tail = TRUE)
pnorm(mean(df$BMI)+2*SD_pop(df$BMI,length(df$BMI)),mean=mean(df$BMI),sd=SD_pop(df$BMI,length(df$BMI)), lower.tail = TRUE)-pnorm(mean(df$BMI)-2*SD_pop(df$BMI,length(df$BMI)),mean=mean(df$BMI),sd=SD_pop(df$BMI,length(df$BMI)), lower.tail = TRUE)
pnorm(mean(df$BMI)+3*SD_pop(df$BMI,length(df$BMI)),mean=mean(df$BMI),sd=SD_pop(df$BMI,length(df$BMI)), lower.tail = TRUE)-pnorm(mean(df$BMI)-3*SD_pop(df$BMI,length(df$BMI)),mean=mean(df$BMI),sd=SD_pop(df$BMI,length(df$BMI)), lower.tail = TRUE)
```
#f.	Create a standard normal distribution and plot the z-scores as a histogram
```{r}
#generate a simulated normal distribution curve  
mean<-mean(df$BMI);sd=SD_pop(df$BMI,length(df$BMI))
x <- seq(-10, 10, by=0.1)
plot(df$BMI,dnorm(df$BMI, mean = mean, sd = sd, log = FALSE))
```

```{r}
hist(z_scores)
```



#Calculate probability for the following examples
#i.	4 individuals enter a clinic, what’s the p that exactly 3 are diabetic (prevalence 10%)
```{r}
dbinom(3,size=4,0.1)
```

#ii.	12 individuals enter a hair salon, what’s the p that exactly 4 have alopecia (prevalence 20% in the studied population)
```{r}
dbinom(4,size=12,0.2)
```

#iii.	15 individuals enter a dentist office, what’s the p that exactly 3 have dental cavity (prevalence 30% in the studied population)
```{r}
dbinom(3,size=15,0.3)
```

#iv.	15 individuals enter a dentist office, what’s the p that at least 3 have dental cavity (prevalence 30% in the studied population)
```{r}
1-(dbinom(0, size=15, prob=0.3)  
     + dbinom(1, size=15, prob=0.3)  
     + dbinom(2, size=15, prob=0.3)
      + dbinom(3,size=15,prob=0.3))
```

#v.	A die is rolled once. What is the probability that the number rolled is even?
```{r}
dbinom(2,size=1,prob = 1/6) + dbinom(4,size=1,prob = 1/6) + dbinom(6,size=1,prob = 1/6)
```

#vi.	A random number is selected between 5 and 15, inclusive. What is the probability that the number selected is less than 10?
```{r}
dbinom(5,size=1,prob=1/11)+ dbinom(6,size=1,prob=1/11)+dbinom(7,size=1,prob=1/11)+dbinom(8,size=1,prob=1/11)+dbinom(9,size=1,prob=1/11)
```

#vii.	If a coin is flipped 5 times, what is the probability of getting exactly 3 heads?
```{r}
dbinom(3,size=5,prob=1/2)
```

#viii.	A test taker guesses answers on a 10-question true/false quiz. What's the probability that they guess exactly 7 correctly?
```{r}
dbinom(7,size=10,1/2)
```

#ix.	The lifetimes of a certain brand of light bulbs are normally distributed with a mean of 800 hours and a standard deviation of 50 hours. What is the probability that a randomly chosen bulb lasts between 750 and 850 hours?
```{r}
z<- (750 - 800) / 50
pnorm(z)
```

#x.	Scores on a particular exam are normally distributed with a mean of 70 and a standard deviation of 10. What percentage of students scored less than 55?
```{r}
z<-(55-70)/10
pnorm(z)
```





